### 2017-05-10 conclusion.tex
Add files via upload

\begin{bibunit}

\rhead{\thepage}

\chapter{Conclusion}

This study presents the first long-term, time-continuous analysis of urban T\textsubscript{surf} and sUHI that incorporates representative sampling of the three-dimensional urban surface. It is based on hemispheric upwelling longwave radiation measurements that are commonly made as a part of urban energy balance assessment and constitute a hitherto untapped resource for urban T\textsubscript{surf} and sUHI analysis. The analysis contributes to a more complete understanding of the temporal and geometric nature of the urban effect on land T\textsubscript{surf} and sUHI.

\section{Summary of results}

Chapter \ref{paper1} details a hemispherical atmospheric correction method that uses version 4.1 of the MODTRAN radiative transfer code \citep{Berk1987} in conjunction with the Surface-Sensor-Sun Urban Model \citep{Soux2004} to derive hemispherical radiometric surface temperatures from longwave radiation upwelling from the urban surface measured from a downward facing pyrgeometer. The correction method is necessary for two reasons:

\begin{enumerate}
	\item To retrieve T\textsubscript{surf} that is representative of the temperature at which the surface is radiating (known as radiometric T\textsubscript{surf}) by removing effects from absorption and emission by the atmosphere in between the surface and the sensor.
	\item To facilitate intersite analysis of T\textsubscript{surf}, where different ambient conditions, surface-sensor-sun geometries, and instrument types can yield differential atmospheric effects on the remote sensed TIR signal.
\end{enumerate}

 The correction method was developed to be readily adaptable to other sensor types and study sites by accounting for complex three-dimensional surface geometry and non uniform spectral sensor response.

Analysis of an eight months of corrected T\textsubscript{hem, r} shows that atmospheric effects on inferred T\textsubscript{hem} are large relative to typical sUHI magnitudes and show significant diurnal and seasonal variation. Relationships between correction magnitude and meteorological variables are complex. Correction magnitudes show strong positive relationships with $\Delta$T\textsubscript{surf-air} and incoming solar radiation. Conversely, although water vapor is the main absorber of broadband TIR radiation, it does not exert significant influence on correction magnitudes, as atmospheric transmittance shows little variation at typical humidities. To facilitate application of the method, a parameterization was developed incorporating a small number of generalizable radiative transfer simulations (included in Appendix \ref{appendixa}) and information about site geometry to retrieve radiometric hemispherical T\textsubscript{surf} en masse using measurements of upwelling longwave, humidity, and T\textsubscript{air}.

In addition, Chapter \ref{paper1} explores the effect of surface geometry on remote sensed urban T\textsubscript{surf} for three representations of the urban surface (complete, plan, and hemispherical). Analysis shows significant overestimation (underestimation) of T\textsubscript{comp} by T\textsubscript{plan} by day (night), particularly under clear sky conditions. By day, T\textsubscript{hem} overestimates T\textsubscript{comp}, but to a lesser degree than T\textsubscript{plan}. Daytime overestimation of T\textsubscript{comp} by T\textsubscript{hem} is more consistent under a wide variety of synoptic conditions, whereas overestimation by T\textsubscript{plan} fluctuates significantly, resulting in poor performance under clear-sky "satellite friendly" conditions. As these measures make up the bulk of the satellite urban T\textsubscript{surf} record, our understanding of urban T\textsubscript{surf} likely includes a significant warm bias. These biases are reflected in sUHI analysis in Chapter \ref{paper2}.

Chapter \ref{paper2} includes an analysis of two long term, time continuous climatologies of hemispherical sUHI, one in Basel, Switzerland and another in Vancouver, Canada. sUHI for both study sites shows significant diurnal, day-to-day, and seasonal variability. sUHI is highly dependent on meteorological conditions, particularly accumulated solar radiation over a day and wind velocity. However, as sUHI is the result of differences in heating and cooling rates between the city and its non urban surroundings, its relationship to meteorological controls is complex. Analysis of sUHI development under a range of mid-latitude synoptic conditions lead to the development of the first observationally supported generalizable model for summertime sUHI development presented in section \ref{gen}.

Analysis of the directional dependence of sUHI is also included in Chapter \ref{paper2}. Results are similar to those found in Chapter \ref{paper1}, with significant daytime overestimation of sUHI\textsubscript{comp} by sUHI{plan}, particularly during hot clear sky conditions. Not only did directional biases in remote sensed sUHI change observed magnitudes, these biases also changed the diurnal character of sUHI, modifying patterns of sUHI in the hours during near sunset. Thus, satellite based assessments of the diurnal pattern of sUHI may lead to misrepresentations of its 'true' diurnal character.

\section{Limitations and future work}

%The convoluted and complex nature of the urban surface makes remote sensing urban T\textsubscript{surf} difficult. Although the method detailed in this study provides T\textsubscript{surf} that is time continuous and more representative of the complete urban T\textsubscript{surf} than conventional remote sensing methods - particularly under clear sky conditions - it is not without some limitations.

The complex nature of the urban surface introduces myriad shortcomings and biases into conventional remote sensing of urban thermal climates - this method is no exception. Sensor placement sensitivity tests in Section \ref{placement} corroborate findings in \citet{Roberts2010,Adderley2015} and indicate that longwave radiation upwelling from the urban surface as measured from a downward facing radiometer is spatially variant when measured from heights below approximately three to five times mean building height. Most pyrgeometers (including those used in this study) are mounted below this height and are subject to a bias towards facets that make up a disproportionately large fraction of the sensor FOV. For sensors mounted at less than optimal heights, the effect of this bias on derived T\textsubscript{hem, r} is largely dependent on sensor placement relative to rooftop facets, with better performance further from positions over rooftops. Pyrgeometers at the Sperrstrasse and Sunset tower sites are placed far enough from rooftops to render this bias relatively small. However, future studies should strive for pyrgeometer heights that are more optimal not only to yield more geometrically representative T\textsubscript{hem, r}, but also to ensure irradiances are representative of the complete urban longwave emission in energy balance assessments.

As the method is fixed and ground based, it is not practical for large scale spatially extensive analysis of sUHI. This limits the method to mid- and low-rise neighborhoods. However, the concept can be adapted to retrieve spatiotemporally continuous urban T\textsubscript{hem, r} from mobile measurements of upwelling longwave irradiances. \citet{Sugawara2006} used near-ground, helicopter-mounted, pyrgeometers and narrow to quantify effects from urban effective anisotropy on upwelling longwave radiation, but did not incorporate a robust atmospheric correction routine to retrieve T\textsubscript{hem, r} and remove spatiotemporally variant atmospheric effects. Inherent complexity in determining the distance to the surface for a moving sensor "seeing" a complex, three-dimensional surface make sensor view modeling difficult. Ray-tracing methods similar to those used in \citet{Ceamanos2017} could provide a solution, but must be modified to return path lengths in addition to seen and unseen points. In addition, sensitivity to ambient conditions (particularly the T\textsubscript{air} profile) will make accurate atmospheric correction computationally expensive in mobile applications of the method. Notwithstanding evaluations of method performance in such applications. If these problems are solved, mobile application of the method may be useful for determine as hemispherical instrumentation can be run simultaneously with near-ground and satellite based narrow-FOV to further understanding of effective anisotropy on TIR remote sensing of urban areas. Such analysis could be used to develop parameterizations to modify satellite retrieved urban T\textsubscript{surf} to better represent the complete urban T\textsubscript{surf}.

Currently, little is known about. The simple fact that issues raised in have yet to be addressed should prompt diverse assessment of urban T\textsubscript{surf} via a range of sensors operating at wide range of spatiotemporal scales is necessary to foster better understand the true geometric, temporal, and spatial effects of urban areas on land T\textsubscript{surf}. While satellite assessment of sUHI and urban T\textsubscript{surf} has been instrumental in elucidating the spatial character of urban effects on T\textsubscript{surf}, several of the questions posed nearly 30 years ago in \citet{Roth1989} remain unanswered. This should prompt not only critical assessment of the urban T\textsubscript{surf} record, but also the development and use of novel methods to fill knowledge gaps, quantify biases, and uncover methodological shortcomings. This method, and in particular its parameterization, is intended to aid in such an endeavor. According to the Urban Flux Network (\href{http://fluxnet.fluxdata.org/}{Urbanflux}) there are 26 active urban flux towers, with many other existing inactive towers, in addition to datasets not indexed under the Urbanflux umbrella for which this method can be used for urban T\textsubscript{surf} and sUHI analysis. These span several continents, a range of climate zones and socioeconomic strata, with many operational over very long (5+ years) time scales. The application of this method over these datasets, and as a component of future energy balance assessments, could provide a wealth of potentially important analysis within and outside the purview of urban climatology.

\section{Final remarks}

%Broadly, the method, its parameterization, and the analysis presented in this study not only provides for novel analysis of the urban effect on T\textsubscript{surf}, but also aims to prompt further critical assessment of the urban T\textsubscript{surf} and the sUHI record.

Broadly, the method, its parameterization, and the analysis presented in this study was conducted to provide for novel analyses of the urban effect on T\textsubscript{surf} and to prompt further critical assessment of the urban T\textsubscript{surf} and sUHI record.  work towards addressing methodological and conceptual questions outlined in \citet{Roth1989}. Indeed, if similar conceptual flaws and shortcomings to those found for the aUHI in \citet{Stewart2011} are likely in sUHI analysis, quantification of instrument and methodological biases and shortcomings in remote sensing of urban T\textsubscript{surf} and the introduction of new methods to overcome these shortcomings may bring us closer to uncovering the true geometric, temporal, and spatial effects of cities on land surface temperature and on the climate at large.

\cleardoublepage 
\phantomsection  
\renewcommand*{\bibname}{References}
\addcontentsline{toc}{section}{\textbf{References}}

\putbib
\end{bibunit}